---
title: "Plant-plant interactions in varietal mixtures directly modulate wheat susceptibility to Septoria tritici blotch"
author: "Laura MATHIEU"
date: "2024"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Load packages

```{r packages}

library(ggplot2)     # For data visualization
library(car)         # For data manipulation
library(dplyr)       # For data manipulation
library(tidyr)       # For data manipulation
library(ARTool)      # For Aligned Rank Transform and ANOVA
library(emmeans)     # For corrected means
library(svglite)     # For saving plots as SVG

```



# Load data for durum wheat - STB



```{r data - DW, results="hide"}

# Remove all objects from the current workspace
remove(list=ls())

# Set the working directory to the specified path
setwd("D:/Mes Donnees/LAURA/Redaction/1. Matrices/Matrices - v2/Data & script")

# Load the dataset from a CSV file
data_DW <- read.table("Durum_wheat_STB.csv", sep = ";", dec = ".", head = TRUE)

# Convert specific columns to appropriate data types
data_DW$Focal <- as.factor(data_DW$Focal)         # Convert 'Focal' to a factor
data_DW$Neighbor <- as.factor(data_DW$Neighbor)   # Convert 'Neighbor' to a factor
data_DW$Pair <- as.factor(data_DW$Pair)           # Convert 'Pair' to a factor
data_DW$Condition <- as.factor(data_DW$Condition) # Convert 'Condition' to a factor
data_DW$N <- as.numeric(data_DW$N)                # Convert 'N' to numeric
data_DW$PN <- as.numeric(data_DW$PN)              # Convert 'PN' to numeric

# Display the structure of the dataset to verify the changes
str(data_DW)

```


# Necrosis analysis


## Mean pairwise comparisons

```{r mean pairwise comparisons - DW - N}

# Fit an aligned rank transform (ART) model
model <- art(N ~ Pair, data = data_DW)

# Perform ANOVA on the ART model
anova(model)

# Conduct pairwise comparisons using the ART model
sign <- art.con(model, "Pair", adjust = "none")

# Increase the maximum number of printed lines to display all results
options(max.print=5000)

# Convert the pairwise comparisons to a data frame
sign_table <- data.frame(sign)

# Apply Bonferroni correction: filter results with p-value < 0.0125 (0.1/8)
sign_table2 <- sign_table[sign_table$p.value < 0.0125,]

# Annotated pairwise comparisons with significant results after Bonferroni correction:
# anv_anv - anv_ato : p-value = 1.646967e-03, adjusted p-value = 0.01317574, significance = *
# ato_anv - ato_ato : p-value = 4.130543e-04, adjusted p-value = 0.003304434, significance = **
# lib_lib - lib_scu : p-value = 4.820292e-03, adjusted p-value = 0.03856234, significance = *
# obe_obe - obe_sur : p-value = 1.114369e-02, adjusted p-value = 0.08914952, significance = .
# scu_lib - scu_scu : p-value = 2.751775e-03, adjusted p-value = 0.0220142, significance = *
# sur_cul - sur_sur : p-value = 2.158203e-04, adjusted p-value = 0.001726562, significance = **
# sur_lib - sur_sur : p-value = 1.023059e-05, adjusted p-value = 8.184472e-05, significance = ***
# voi_scu - voi_voi : p-value = 2.582799e-03, adjusted p-value = 0.02066239, significance = *

# 8 mixtures = 9.88% mixtures showing NMS

```

## Create mean heatmap for pairwise comparisons

```{r mean pairwise heatmap - DW - N , include=FALSE}

# Generate estimated marginal means (least squares means)
emm <- emmeans(artlm(model, "Pair"), ~ Pair, pbktest.limit = 200000)
leastsquare <- as.data.frame(emm)

# Merge the estimated marginal means with the original dataset
data_DW_2 <- merge(leastsquare, data_DW, by = "Pair")

# Initialize a data frame to store the results
Pair <- unique(data_DW_2$Pair)
Paires <- data.frame()
compt <- 1

# Calculate the mean 'emmean' for each pair
for(i in Pair) {
  sub <- subset(data_DW_2, Pair == i)
  x <- mean(sub$emmean)
  Paires[compt,1] <- i
  Paires[compt,2] <- x
  f <- unique(sub$Focal)
  n <- unique(sub$Neighbor)
  Paires[compt,3] <- f
  Paires[compt,4] <- n
  compt <- compt + 1
}

# Rename the columns of the data frame
colnames(Paires) <- c('Pair', 'emmean', 'Focal', 'Neighbor')

# Initialize data frames for storing normalized values and pairs
Focal <- unique(Paires$Focal)
liste <- data.frame(nrow = 1:9)
liste$nrow <- as.character(liste$nrow)
liste_pair <- data.frame(nrow = 1:9)
liste_pair$nrow <- as.character(liste_pair$nrow)

# Normalize the emmeans by the mean of each focal pair
for(i in Focal) {
  pair_sub <- subset(Paires, Focal == i)
  
  pair_sub_ii <- subset(pair_sub, Neighbor == i)
  x <- pair_sub_ii$emmean
  
  y <- pair_sub$emmean
  norm <- pair_sub$emmean / x
  norm <- as.data.frame(norm)
  norm[,2] <- row.names(norm)
  colnames(norm) <- c('norm', 'nrow')
  liste <- liste %>% full_join(norm, by = "nrow")
  
  pair <- pair_sub$Pair
  pair <- as.data.frame(pair)
  pair[,2] <- row.names(pair)
  colnames(pair) <- c('Pair', 'nrow')
  liste_pair <- liste_pair %>% full_join(pair, by = "nrow")
}

# Process the normalized values and pairs
liste <- liste[,2:10]
colnames(liste) <- 1:9
list <- as.list(liste)
list_num <- data.frame(unlist(list))
list_ssNA <- subset(list_num, unlist.list.!="NA")

liste_pair <- liste_pair[,2:10]
colnames(liste_pair) <- 1:9
list_pair <- as.list(liste_pair)
list_pair <- data.frame(unlist(list_pair))
list_pair_ssNA <- subset(list_pair, unlist.list_pair.!="NA")

# Merge the processed values with the original pairs
y_pur <- merge(list_ssNA, list_pair_ssNA, by = 0)
colnames(y_pur) <- c('nrow','y_pur','Pair')
Pair <- merge(y_pur, Paires, by = "Pair")

# Reorder the levels of the 'Focal' factor
Pair$Focal <- factor(Pair$Focal, levels = rev(unique(Paires$Focal)))

# Create a mean heatmap using ggplot2
heatmap_y_mix_pure_mean_pure <- ggplot(Pair, aes(x = Neighbor, y = Focal)) + 
  geom_tile(aes(fill = y_pur), colour = "black", size = 1) +
  scale_fill_gradient2(low = "darkblue", high = "red", mid = "white", midpoint = 1) +
  theme_minimal() +
  geom_text(x = 2, y = 8.8, label = "*", col = "black", size = 10) +
  geom_text(x = 1, y = 7.8, label = "**", col = "black", size = 10) +
  geom_text(x = 7, y = 5.8, label = "*", col = "black", size = 10) +
  geom_text(x = 8, y = 5.3, label = ".", col = "black", size = 10) +
  geom_text(x = 4, y = 2.8, label = "*", col = "black", size = 10) +
  geom_text(x = 3, y = 1.8, label = "**", col = "black", size = 10) +
  geom_text(x = 4, y = 1.8, label = "***", col = "black", size = 10) +
  geom_text(x = 7, y = 0.8, label = "*", col = "black", size = 10) +
  ylab("Focal") +
  xlab("Neighbor") +
  ggtitle("Necrosis surface / leaf surface of the focal plant (N/L)") +
  labs(fill = "Mean(Mix) / Mean(Pure)") +
  theme(plot.title = element_text(color = "black", size = 15, face = "italic", hjust = 0.5),
        legend.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 12, face = "bold", angle = 90, hjust = 1),
        axis.title.x = element_text(color = "black", size = 12, face = "bold"),
        strip.text.x = element_text(color = "black", size = 12, face = "bold"),
        axis.text.y = element_text(color = "black", size = 12, face = "bold", angle = 0, hjust = 1),
        axis.title.y = element_text(color = "black", size = 12, face = "bold", angle = 90, hjust = 1))

# Display the heatmap
heatmap_y_mix_pure_mean_pure

# Save the heatmap as an SVG file
ggsave(file="Matrice_NonParam_DW_Necrosis.svg", plot=heatmap_y_mix_pure_mean_pure, width=9, height=9)

```

## Variance pairwise comparisons

```{r variance pairwise comparisons - DW - N}

# Initialize variables
Pair = unique(data_DW$Pair)
Paires = data.frame()
compt = 1

# Calculate variances for each pair and store in Paires dataframe
for (i in Pair) { 
  sub = subset(data_DW, Pair == i)
  x = var(sub$N)
  print(x)
  Paires[compt, 1] = i
  Paires[compt, 2] = x
  f = unique(sub$Focal)
  n = unique(sub$Neighbor)
  Paires[compt, 3] = f
  Paires[compt, 4] = n
  compt = compt + 1
}

# Rename columns for clarity
colnames(Paires) = c('Pair', 'var', 'Focal', 'Neighbor')

# Initialize dataframes for normalized values and pairs
Focal = unique(Paires$Focal)
liste = data.frame(nrow = 1:9)
liste$nrow = as.character(liste$nrow)
liste_pair = data.frame(nrow = 1:9)
liste_pair$nrow = as.character(liste_pair$nrow)

# Normalize the variances by the mean of each focal pair
for (i in Focal) {
  pair_sub = subset(Paires, Focal == i)
  
  pair_sub_ii = subset(pair_sub, Neighbor == i)
  x = pair_sub_ii$var
  print(x)
  
  y = pair_sub$var
  print(y)
  norm = pair_sub$var / x
  norm = as.data.frame(norm)
  norm[, 2] = row.names(norm)
  colnames(norm) = c('norm', 'nrow')
  print(norm)
  liste <- liste %>% full_join(norm, by = "nrow")
  
  pair = pair_sub$Pair
  pair = as.data.frame(pair)
  pair[, 2] = row.names(pair)
  colnames(pair) = c('Pair', 'nrow')
  liste_pair <- liste_pair %>% full_join(pair, by = "nrow")
}

# Process the normalized values and pairs
liste = liste[, 2:10]
colnames(liste) = 1:9
list = as.list(liste)
list_num = data.frame(unlist(list))
list_ssNA = subset(list_num, unlist.list.!="NA")

liste_pair = liste_pair[, 2:10]
colnames(liste_pair) = 1:9
list_pair = as.list(liste_pair)
list_pair = data.frame(unlist(list_pair))
list_pair_ssNA = subset(list_pair, unlist.list_pair.!="NA")

# Merge the processed values with the original pairs
y_pur = merge(list_ssNA, list_pair_ssNA, by = 0)
colnames(y_pur) = c('nrow', 'y_pur', 'Pair')
Pair = merge(y_pur, Paires, by = "Pair")

# Perform pairwise variance tests
pvals_var_all = data.frame()

Focal = unique(Paires$Focal)
Neighbor = unique(Paires$Neighbor)

for (i in Focal) {
  pair_sub = subset(data_DW, Focal == i)
  
  for (j in Neighbor) {
    # Execute the loop only if j is different from i
    if (j != i) {
      pair_sub_ii = subset(pair_sub, Neighbor == i | Neighbor == j)
      
      levene = leveneTest(N ~ Condition, data = pair_sub_ii)
      fligner = fligner.test(N ~ Condition, data = pair_sub_ii)

      pvals_var = data.frame()
      pvals_var[1, 1] = i
      pvals_var[1, 2] = j
      pvals_var[1, 3] = levene$`Pr(>F)`[1]
      pvals_var[1, 4] = fligner$p.value

      colnames(pvals_var) = c('Focal', 'Neighbor', 'pval_Levene', 'pval_Fligner')
      pvals_var_all = rbind(pvals_var_all, pvals_var)
    }
  }
}

# Adjust p-values for multiple comparisons (Bonferroni correction)
pvals_var_all$pval_Levene = pvals_var_all$pval_Levene * 8
pvals_var_all$pval_Fligner = pvals_var_all$pval_Fligner * 8
pvals_var_all = pvals_var_all[order(pvals_var_all$pval_Fligner), ]

# Annotated pairwise comparisons with significant results after Bonferroni correction:
# cul - lib:  pval_Levene = 7.397894e-07, pval_Fligner = 0.0001365887, significance = ***
# cul - scu:  pval_Levene = 5.654614e-05, pval_Fligner = 0.0023446940, significance = **
# cul - obe:  pval_Levene = 1.166195e-03, pval_Fligner = 0.0118686547, significance = *
# voi - cul:  pval_Levene = 1.534590e-02, pval_Fligner = 0.0120647100, significance = *
# scu - pes:  pval_Levene = 2.014455e-02, pval_Fligner = 0.0219987636, significance = *
# lib - sur:  pval_Levene = 5.872274e-02, pval_Fligner = 0.0665102202, significance = .

# Summary: 6 mixtures (7.41% of mixtures) have a variance different from pure

```

## Variance heatmap of pairwise comparisons

```{r mean pairwise heatmap - DW - N , include=FALSE}

# Reorder the levels of the 'Focal' factor
Pair$Focal = factor(Pair$Focal, levels = rev(sort(unique(Pair$Focal))))

# Create a heatmap using ggplot2
heatmap_y_mix_pure_var <- ggplot(Pair, aes(x = Neighbor, y = Focal)) + 
  geom_tile(aes(fill = y_pur), colour = "black", size = 1) +
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "red", midpoint = 1, na.value = "#A52A2A") +
  theme_minimal() +
  
  geom_text(x = 4, y = 6.8, label = "***", col = "black", size = 10) +
  geom_text(x = 7, y = 6.8, label = "**", col = "black", size = 10) +
  geom_text(x = 5, y = 6.8, label = "*", col = "black", size = 10) +
  geom_text(x = 3, y = 0.8, label = "*", col = "black", size = 10) +
  geom_text(x = 6, y = 2.8, label = "*", col = "black", size = 10) +
  geom_text(x = 8, y = 6.3, label = ".", col = "black", size = 10) +
  
  ylab("Focal") +
  xlab("Neighbor") +
  ggtitle("Necrosis surface / leaf surface of the focal plant (N/L)") +
  labs(fill = "Var(Mix) / Var(Pure)") +
  theme(plot.title = element_text(color = "black", size = 15, face = "italic", hjust = 0.5), # Title formatting
        legend.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 12, face = "bold", angle = 90, hjust = 1),
        axis.title.x = element_text(color = "black", size = 12, face = "bold"),
        strip.text.x = element_text(color = "black", size = 12, face = "bold"),
        axis.text.y = element_text(color = "black", size = 12, face = "bold", angle = 0, hjust = 1),
        axis.title.y = element_text(color = "black", size = 12, face = "bold", angle = 90, hjust = 1)
  )

# Display the heatmap
heatmap_y_mix_pure_var

# Save the heatmap as an SVG file
library(svglite)
ggsave(file = "Matrice_NonParam_var_DW_Necrosis.svg", plot = heatmap_y_mix_pure_var, width = 9, height = 9)

```

## Variance explained by genetic effects

```{r DGE-IGE - DW - N, include=FALSE}

# Fit the aligned rank transform model to the data
model = art(N ~ Focal * Neighbor, data = data_DW_2)

# Perform ANOVA on the fitted model
Result = anova(model)

# Calculate partial eta squared for each term in the ANOVA table
Result$part.eta.sq = with(Result, `Sum Sq` / (`Sum Sq` + `Sum Sq.res`))

# Display the ANOVA result with partial eta squared
Result

# Perform ANOVA on the fitted model again for redundancy (to show another way to calculate partial eta squared)
Result2 = anova(model)

# Calculate partial eta squared using an alternative method
Result2$part.eta.sq = with(Result2, `F value` * `Df` / (`F value` * `Df` + `Df.res`))

# Display the second ANOVA result with partial eta squared
Result2

```

## Boxplots mix-pur with variances & means

```{r mean - variances - DW - N}

# Conduct Wilcoxon test & fligner test for N by condition
wilcox.test(N~Condition, data=data_DW)
fligner.test(N~Condition, data=data_DW)

# Calculate variances for mix and pure conditions
Mix <- subset(data_DW, data_DW$Condition == "Mix")
Mix <- Mix$N
var(Mix)
Pure <- subset(data_DW, data_DW$Condition == "Pure")
Pure <- Pure$N
var(Pure)

# Plotting
violin <- ggplot(data_DW_2, aes(x = factor(Condition, level = c('Pure', 'Mix')), y=N)) + geom_violin(trim=FALSE)
plot <- violin + geom_boxplot(width=0.3, col="#9cb264") + stat_summary(fun=mean, geom="point", shape=23, size=5, color="#7a3776")+
  geom_text(x=1.5, y=145, label="ns", col="#7a3776", size = 10)+
  geom_text(x=1, y=145, label="μ(pure) = 66.1", col="#7a3776", size = 10)+
  geom_text(x=2, y=145, label="μ(mix) = 66.27", col="#7a3776", size = 10)+
  geom_text(x=1.5, y=135, label="*", col="#9cb264", size = 20)+
  geom_text(x=1, y=135, label="var(pure) = 1190.96", col="#9cb264", size = 10)+
  geom_text(x=2, y=135, label="var(mix) = 1142.37", col="#9cb264", size = 10)+
  ylab("Necrosis surface / leaf surface (N/L)") +
  ggtitle("Durum wheat")+
  theme(plot.title =element_text(color = "black", size = 40, face = "italic",hjust =0.5),
    axis.text.x = element_text(color = "black", size = 40, face = "bold",angle = 0, hjust = ),
  axis.title.x = element_text(color = "black", size = 40, face = "bold"), 
  strip.text.x = element_text(color = "black",size=40, face = "bold"),
  axis.text.y = element_text(color = "black", size = 40, face = "bold",angle = 0, hjust = ),
  axis.title.y = element_text(color = "black", size = 40, face = "bold",angle = 90, hjust = ))+
  ylim(-25,150)
print(plot)

# Save the plot as an SVG file
ggsave(file="MixPur_brut_mean_var_DW_Necrosis.svg", plot=plot, width=15, height=15)

```

## Variances vs. mean analysis 

```{r var ~ mean - DW - N, include=FALSE}

# Initialize variables and data frame
pair = unique(data_DW$Pair)
sf.N.L_pair <- data.frame()
n = 1

# Loop through each modality to calculate variance and mean of necrosis surface
for (i in pair) {
  symptom = subset(data_DW, Pair == i)  # Subset data for the current pair
  v = var(symptom$N)  # Calculate variance of necrosis surface
  m = mean(symptom$N)  # Calculate mean of necrosis surface
  print(v)
  print(m)
  new = c(i, v, m)
  print(new)
  # Add the results to the data frame
  sf.N.L_pair[nrow(sf.N.L_pair) + 1, 1:3] <- c(i, v, m)
  n = n + 1
}

# Rename columns
colnames(sf.N.L_pair) <- c('Pair', 'Var', 'Mean')
sf.N.L_pair[,2] <- as.numeric(sf.N.L_pair[,2])
sf.N.L_pair[,3] <- as.numeric(sf.N.L_pair[,3])
# Order the data frame by mean
sf.N.L_pair <- sf.N.L_pair[order(sf.N.L_pair[,3]),]

# Merge the calculated variance and mean back to the original data
data_DW_3 <- merge(data_DW, sf.N.L_pair)
# Order the merged data by the calculated mean
data_DW_3 <- data_DW_3[order(data_DW_3[,9]),]

# Create a boxplot of necrosis surface by pair, ordered by mean
plot_var <- ggplot(data_DW_3, aes(x = reorder(Pair, Mean), y = N)) + 
  geom_boxplot(fill = "red", outlier.colour = "black", outlier.shape = 16,
               outlier.size = 2, notch = FALSE)

# Display the boxplot
plot_var

# Create a scatter plot of mean vs. variance of necrosis surface
plot <- ggplot(sf.N.L_pair, aes(x = Mean, y = Var)) + 
  geom_point() +
  geom_smooth(color = "darkred", fill = "red") +
  xlab("Average necrosis surface of focal plants") +
  ylab("Necrosis variance of focal plants") +
  theme(
    axis.text.x = element_text(color = "black", size = 20, face = "bold", angle = 0, hjust = ),
    axis.title.x = element_text(color = "black", size = 20, face = "bold"), 
    strip.text.x = element_text(color = "black", size = 20, face = "bold"),
    axis.text.y = element_text(color = "black", size = 20, face = "bold", angle = 0, hjust = ),
    axis.title.y = element_text(color = "black", size = 20, face = "bold", angle = 90, hjust = )
  )

# Display the scatter plot
plot

# Save the scatter plot to a file
ggsave(file = "Corr_mean_var__DW_Necrosis.svg", plot = plot, width = 7, height = 7)

```


# Pycnidia analysis


## Mean pairwise comparisons

```{r mean pairwise - DW - P}

# Remove NA values from data_DW$PN
data_DW <- data_DW[!is.na(data_DW$PN),]

# Fit an aligned rank transform (ART) model
model = art(PN ~ Pair, data = data_DW)

# Perform ANOVA on the ART model
anova(model)

# Conduct pairwise comparisons using the ART model
sign <- art.con(model, "Pair", adjust = "none")

# Increase the maximum number of printed lines to display all results
options(max.print=5000)

# Convert the pairwise comparisons to a data frame
sign_table <- data.frame(sign)

# Apply Bonferroni correction: filter results with p-value < 0.0125 (0.1/8)
sign_table2 <- sign_table[sign_table$p.value < 0.0125,]

# Annotated pairwise comparisons with significant results after Bonferroni correction:
# anv_anv - anv_ato : p-value = 9.128641e-07, adjusted p-value = 7.302913e-06, significance = ***
# anv_anv - anv_obe : p-value = 6.471299e-06, adjusted p-value = 5.177039e-05, significance = ***
# ato_ato - ato_scu : p-value = 2.149703e-03, adjusted p-value = 0.01719762, significance = *
# cul_anv - cul_cul : p-value = 2.024464e-03, adjusted p-value = 0.01619571, significance = *
# cul_cul - cul_lib : p-value = 5.234397e-04, adjusted p-value = 0.004187518, significance = **
# cul_cul - cul_pes : p-value = 1.169339e-02, adjusted p-value = 0.09354712, significance = .
# pes_lib - pes_pes : p-value = 4.843422e-03, adjusted p-value = 0.03874738, significance = *
# pes_obe - pes_pes : p-value = 5.417145e-03, adjusted p-value = 0.04333716, significance = *
# pes_pes - pes_voi : p-value = 1.109630e-02, adjusted p-value = 0.0887704, significance = .
# scu_scu - scu_sur : p-value = 7.659616e-03, adjusted p-value = 0.06127693, significance = .
# sur_scu - sur_sur : p-value = 6.032668e-03, adjusted p-value = 0.04826134, significance = *

# 11 mixtures = 13.58% of mixtures showing significant differences

```

## Create mean heatmap for pairwise comparisons

```{r mean pairwise heatmap - DW - P, include=FALSE}

# Generate estimated marginal means (least squares means)
emm <- emmeans(artlm(model, "Pair"), ~ Pair, pbktest.limit = 200000)

# Convert the estimated marginal means to a data frame
leastsquare <- as.data.frame(emm)

# Merge the estimated marginal means with the original dataset
data_DW_2 <- merge(leastsquare, data_DW, by = "Pair")

# Initialize a data frame to store the results
Pair <- unique(data_DW_2$Pair)
Paires <- data.frame()
compt <- 1

# Calculate the mean 'emmean' for each pair
for(i in Pair) {
  sub <- subset(data_DW_2, Pair == i)
  x <- mean(sub$emmean)
  Paires[compt, 1] <- i
  Paires[compt, 2] <- x
  f <- unique(sub$Focal)
  n <- unique(sub$Neighbor)
  Paires[compt, 3] <- f
  Paires[compt, 4] <- n
  compt <- compt + 1
}

# Rename the columns of the data frame
colnames(Paires) <- c('Pair', 'emmean', 'Focal', 'Neighbor')

# Initialize data frames for storing normalized values and pairs
Focal <- unique(Paires$Focal)
liste <- data.frame(nrow = 1:9)
liste$nrow <- as.character(liste$nrow)
liste_pair <- data.frame(nrow = 1:9)
liste_pair$nrow <- as.character(liste_pair$nrow)

# Normalize the emmeans by the mean of each focal pair
for(i in Focal) {
  pair_sub <- subset(Paires, Focal == i)
  
  pair_sub_ii <- subset(pair_sub, Neighbor == i)
  x <- pair_sub_ii$emmean
  
  y <- pair_sub$emmean
  norm <- pair_sub$emmean / x
  norm <- as.data.frame(norm)
  norm[, 2] <- row.names(norm)
  colnames(norm) <- c('norm', 'nrow')
  liste <- liste %>% full_join(norm, by = "nrow")
  
  pair <- pair_sub$Pair
  pair <- as.data.frame(pair)
  pair[, 2] <- row.names(pair)
  colnames(pair) <- c('Pair', 'nrow')
  liste_pair <- liste_pair %>% full_join(pair, by = "nrow")
}

# Process the normalized values and pairs
liste <- liste[, 2:10]
colnames(liste) <- 1:9
list <- as.list(liste)
list_num <- data.frame(unlist(list))
list_ssNA <- subset(list_num, unlist.list.!="NA")

liste_pair <- liste_pair[, 2:10]
colnames(liste_pair) <- 1:9
list_pair <- as.list(liste_pair)
list_pair <- data.frame(unlist(list_pair))
list_pair_ssNA <- subset(list_pair, unlist.list_pair.!="NA")

# Merge the processed values with the original pairs
y_pur <- merge(list_ssNA, list_pair_ssNA, by = 0)
colnames(y_pur) <- c('nrow', 'y_pur', 'Pair')
Pair <- merge(y_pur, Paires, by = "Pair")

# Reorder the levels of the 'Focal' factor
Pair$Focal <- factor(Pair$Focal, levels = rev(unique(Paires$Focal)))

# Create a mean heatmap using ggplot2
heatmap_y_mix_pure_mean_pure <- ggplot(Pair, aes(x = Neighbor, y = Focal)) + 
  geom_tile(aes(fill = y_pur), colour = "black", size = 1) +
  scale_fill_gradient2(low = "darkblue", high = "red", mid = "white", midpoint = 1) +
  theme_minimal() +
  geom_text(x = 2, y = 8.8, label = "***", col = "black", size = 10) +
  geom_text(x = 5, y = 8.8, label = "***", col = "black", size = 10) +
  geom_text(x = 7, y = 7.8, label = "*", col = "black", size = 10) +
  geom_text(x = 1, y = 6.8, label = "*", col = "black", size = 10) +
  geom_text(x = 4, y = 6.8, label = "**", col = "black", size = 10) +
  geom_text(x = 6, y = 7.3, label = ".", col = "black", size = 10) +
  geom_text(x = 4, y = 3.8, label = "*", col = "black", size = 10) +
  geom_text(x = 5, y = 3.8, label = "*", col = "black", size = 10) +
  geom_text(x = 9, y = 4.3, label = ".", col = "black", size = 10) +
  geom_text(x = 8, y = 3.3, label = ".", col = "black", size = 10) +
  geom_text(x = 7, y = 1.8, label = "*", col = "black", size = 10) +
  ylab("Focal") +
  xlab("Neighbor") +
  ggtitle("Pycnidia surface / leaf surface of the focal plant (P/L)") +
  labs(fill = "Mean(Mix) / Mean(Pure)") +
  theme(plot.title = element_text(color = "black", size = 15, face = "italic", hjust = 0.5),
        legend.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 12, face = "bold", angle = 90, hjust = 1),
        axis.title.x = element_text(color = "black", size = 12, face = "bold"),
        strip.text.x = element_text(color = "black", size = 12, face = "bold"),
        axis.text.y = element_text(color = "black", size = 12, face = "bold", angle = 0, hjust = 1),
        axis.title.y = element_text(color = "black", size = 12, face = "bold", angle = 90, hjust = 1))

# Display the heatmap
heatmap_y_mix_pure_mean_pure

# Save the heatmap as SVG file
ggsave(file="Matrice_NonParam_DW_Pycnides.svg", plot=heatmap_y_mix_pure_mean_pure, width=9, height=9)

```

## Variance paiwise comparisons 

```{r variance paiwise comparisons - DW - P}

# Initialize variables
Pair <- unique(data_DW$Pair)
Paires <- data.frame()
compt <- 1

# Calculate variances for each pair and store in Paires dataframe
for (i in Pair) { 
  sub <- subset(data_DW, Pair == i)
  x <- var(sub$PN)
  print(x)
  Paires[compt, 1] <- i
  Paires[compt, 2] <- x
  f <- unique(sub$Focal)
  n <- unique(sub$Neighbor)
  Paires[compt, 3] <- f
  Paires[compt, 4] <- n
  compt <- compt + 1
}

# Rename columns for clarity
colnames(Paires) <- c('Pair', 'var', 'Focal', 'Neighbor')

# Initialize dataframes for normalized values and pairs
Focal <- unique(Paires$Focal)
liste <- data.frame(nrow = 1:9)
liste$nrow <- as.character(liste$nrow)
liste_pair <- data.frame(nrow = 1:9)
liste_pair$nrow <- as.character(liste_pair$nrow)

# Normalize the variances by the mean of each focal pair
for (i in Focal) {
  pair_sub <- subset(Paires, Focal == i)
  
  pair_sub_ii <- subset(pair_sub, Neighbor == i)
  x <- pair_sub_ii$var
  print(x)
  
  y <- pair_sub$var
  print(y)
  norm <- pair_sub$var / x
  norm <- as.data.frame(norm)
  norm[, 2] <- row.names(norm)
  colnames(norm) <- c('norm', 'nrow')
  print(norm)
  liste <- liste %>% full_join(norm, by = "nrow")
  
  pair <- pair_sub$Pair
  pair <- as.data.frame(pair)
  pair[, 2] <- row.names(pair)
  colnames(pair) <- c('Pair', 'nrow')
  liste_pair <- liste_pair %>% full_join(pair, by = "nrow")
}

# Process the normalized values and pairs
liste <- liste[, 2:10]
colnames(liste) <- 1:9
list <- as.list(liste)
list_num <- data.frame(unlist(list))
list_ssNA <- subset(list_num, unlist.list.!="NA")

liste_pair <- liste_pair[, 2:10]
colnames(liste_pair) <- 1:9
list_pair <- as.list(liste_pair)
list_pair <- data.frame(unlist(list_pair))
list_pair_ssNA <- subset(list_pair, unlist.list_pair.!="NA")

# Merge the processed values with the original pairs
y_pur <- merge(list_ssNA, list_pair_ssNA, by = 0)
colnames(y_pur) <- c('nrow', 'y_pur', 'Pair')
Pair <- merge(y_pur, Paires, by = "Pair")

# Perform pairwise variance tests
pvals_var_all <- data.frame()

Focal <- unique(Paires$Focal)
Neighbor <- unique(Paires$Neighbor)

for (i in Focal) {
  pair_sub <- subset(data_DW, Focal == i)
  
  for (j in Neighbor) {
    # Execute the loop only if j is different from i
    if (j != i) {
      pair_sub_ii <- subset(pair_sub, Neighbor == i | Neighbor == j)
      
      levene <- leveneTest(PN ~ Condition, data = pair_sub_ii)
      fligner <- fligner.test(PN ~ Condition, data = pair_sub_ii)

      pvals_var <- data.frame()
      pvals_var[1, 1] <- i
      pvals_var[1, 2] <- j
      pvals_var[1, 3] <- levene$`Pr(>F)`[1]
      pvals_var[1, 4] <- fligner$p.value

      colnames(pvals_var) <- c('Focal', 'Neighbor', 'pval_Levene', 'pval_Fligner')
      pvals_var_all <- rbind(pvals_var_all, pvals_var)
    }
  }
}

# Adjust p-values for multiple comparisons (Bonferroni correction)
pvals_var_all$pval_Levene <- pvals_var_all$pval_Levene * 8
pvals_var_all$pval_Fligner <- pvals_var_all$pval_Fligner * 8
pvals_var_all <- pvals_var_all[order(pvals_var_all$pval_Fligner), ]

# Annotated pairwise comparisons with significant results after Bonferroni correction:
#   ato   sur   0.03723385    0.0008649421    ***
#   sur   scu   0.02069371    0.0021172577    **
#   scu   sur   0.15001322    0.0047978954    **
#   ato   scu   0.15158168    0.0085245977    **
#   ato   voi   0.63510489    0.0103441271    *
#   ato   anv   0.96088780    0.0177670145    *
#   cul   pes   1.73836918    0.0227830824    *
#   sur   cul   0.35855381    0.0267056063    *
#   ato   obe   1.37829767    0.0322808505    *
#   scu   anv   0.76086150    0.0826256500    .
#   ato   cul   4.04637651    0.0976248634    .

# Summary: 11 mixtures (13.58% of mixtures) have a variance different from pure

```

## Variance heatmap of pairwise comparisons

```{r variance pairwise heatmap - DW - P, include=FALSE}

# Replace values > 4 with NA
Pair$y_pur <- ifelse(Pair$y_pur > 4, NA, Pair$y_pur)

# Reorder the levels of the 'Focal' factor
Pair$Focal <- factor(Pair$Focal, levels = rev(sort(unique(Pair$Focal))))

# Create a heatmap using ggplot2
heatmap_y_mix_pure_var <- ggplot(Pair,aes(x=Neighbor,y= Focal)) + 
  geom_tile(aes(fill = y_pur), colour = "black", size = 1)+
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "red", midpoint = 1, na.value = "#A52A2A") +
  theme_minimal()+

  geom_text(x=8, y=7.8, label="***", col="black", size = 10)+
  geom_text(x=7, y=1.8, label="**", col="black", size = 10)+
  geom_text(x=8, y=2.8, label="**", col="black", size = 10)+
  geom_text(x=7, y=7.8, label="**", col="black", size = 10)+
  geom_text(x=9, y=7.8, label="*", col="black", size = 10)+
  geom_text(x=1, y=7.8, label="*", col="black", size = 10)+
  geom_text(x=6, y=6.8, label="*", col="black", size = 10)+
  geom_text(x=3, y=1.8, label="*", col="black", size = 10)+
  geom_text(x=5, y=7.8, label="*", col="black", size = 10)+
  geom_text(x=1, y=3.3, label=".", col="black", size = 10)+
  geom_text(x=3, y=8.3, label=".", col="black", size = 10)+
  
  ylab("Focal") +
  xlab("Neighbor")+
  ggtitle("Pycnidia surface / leaf surface of the focal plant (P/L)")+
  labs(fill="Var(Mix) / Var(Pure)")+
  theme(plot.title =element_text(color = "black", size = 15, face = "italic",hjust =0.5),#mise en forme
        legend.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 12, face = "bold",angle = 90, hjust = ),
        axis.title.x = element_text(color = "black", size = 12, face = "bold"),
        strip.text.x = element_text(color = "black",size=12, face = "bold"),
        axis.text.y = element_text(color = "black", size = 12, face = "bold",angle = 0, hjust = ),
        axis.title.y = element_text(color = "black", size = 12, face = "bold",angle = 90, hjust = )
  )

# Display the heatmap
heatmap_y_mix_pure_var

# Save the heatmap as an SVG file
ggsave(file = "Matrice_NonParam_var_DW_Pycnidia.svg", plot = heatmap_y_mix_pure_var, width = 9, height = 9)

```

## Variance explained by genetic effects

```{r DGE-IGE - DW - P, include=FALSE}

# Fit the aligned rank transform model to the data
model = art(PN ~ Focal * Neighbor, data = data_DW_2)

# Perform ANOVA on the fitted model
Result = anova(model)

# Calculate partial eta squared for each term in the ANOVA table
Result$part.eta.sq = with(Result, `Sum Sq` / (`Sum Sq` + `Sum Sq.res`))

# Display the ANOVA result with partial eta squared
Result

# Perform ANOVA on the fitted model again for redundancy (to show another way to calculate partial eta squared)
Result2 = anova(model)

# Calculate partial eta squared using an alternative method
Result2$part.eta.sq = with(Result2, `F value` * `Df` / (`F value` * `Df` + `Df.res`))

# Display the second ANOVA result with partial eta squared
Result2

```

## Boxplots mix-pur with variances & means

```{r mean - variances - DW - P}

# Conduct Wilcoxon test & fligner test for N by condition
wilcox.test(PN~Condition, data=data_DW)
fligner.test(PN~Condition, data=data_DW)

# Calculate variances for mix and pure conditions
Mix <- subset(data_DW, data_DW$Condition == "Mix")
Mix <- Mix$PN
var(Mix)
Pure <- subset(data_DW, data_DW$Condition == "Pure")
Pure <- Pure$PN
var(Pure)

# Plotting
violin <- ggplot(data_DW_2, aes(x=factor(Condition, level=c("Pure","Mix")), y=PN)) + geom_violin(trim=FALSE)
plot <- violin + geom_boxplot(width=0.3, col="#9cb264") + stat_summary(fun=mean, geom="point", shape=23, size=5, color="#7a3776")+
  geom_text(x=1.5, y=58, label="**", col="#7a3776", size = 20)+
  geom_text(x=1, y=58, label="μ(pure) = 8.91", col="#7a3776", size = 10)+
  geom_text(x=2, y=58, label="μ(mix) = 12.38", col="#7a3776", size = 10)+
  geom_text(x=1.5, y=53, label="***", col="#9cb264", size = 20)+
  geom_text(x=1, y=53, label="var(pure) = 310.38", col="#9cb264", size = 10)+
  geom_text(x=2, y=53, label="var(mix) = 429.27", col="#9cb264", size = 10)+
  ylab("Pycnidia surface / leaf surface (P/L)") +
  ggtitle("Durum wheat")+
  theme(plot.title =element_text(color = "black", size = 40, face = "italic",hjust =0.5),
      axis.text.x = element_text(color = "black", size = 40, face = "bold",angle = 0, hjust = ),
  axis.title.x = element_text(color = "black", size = 40, face = "bold"), 
  strip.text.x = element_text(color = "black",size=40, face = "bold"),
  axis.text.y = element_text(color = "black", size = 40, face = "bold",angle = 0, hjust = ),
  axis.title.y = element_text(color = "black", size = 40, face = "bold",angle = 90, hjust = ))+
  ylim(-25,60)
print(plot)

# Save the plot as an SVG file
ggsave(file="MixPur_brut_mean_var_DW_Pycnidia.svg", plot=plot, width=15, height=15)

```

## Variances vs. mean analysis 

```{r var ~ mean - DW - P, include=FALSE}

# Initialize variables and data frame
pair = unique(data_DW$Pair)
sf.N.L_pair <- data.frame()
n = 1

# Loop through each modality to calculate variance and mean of necrosis surface
for (i in pair) {
  symptom = subset(data_DW, Pair == i)  # Subset data for the current pair
  v = var(symptom$PN)  # Calculate variance of necrosis surface
  m = mean(symptom$PN)  # Calculate mean of necrosis surface
  print(v)
  print(m)
  new = c(i, v, m)
  print(new)
  # Add the results to the data frame
  sf.N.L_pair[nrow(sf.N.L_pair) + 1, 1:3] <- c(i, v, m)
  n = n + 1
}

# Rename columns
colnames(sf.N.L_pair) <- c('Pair', 'Var', 'Mean')
sf.N.L_pair[,2] <- as.numeric(sf.N.L_pair[,2])
sf.N.L_pair[,3] <- as.numeric(sf.N.L_pair[,3])
# Order the data frame by mean
sf.N.L_pair <- sf.N.L_pair[order(sf.N.L_pair[,3]),]

# Merge the calculated variance and mean back to the original data
data_DW_3 <- merge(data_DW, sf.N.L_pair)
# Order the merged data by the calculated mean
data_DW_3 <- data_DW_3[order(data_DW_3[,9]),]

# Create a boxplot of pycnidia by pair, ordered by mean
plot_var <- ggplot(data_DW_3, aes(x = reorder(Pair, Mean), y = PN)) + 
  geom_boxplot(fill = "red", outlier.colour = "black", outlier.shape = 16,
               outlier.size = 2, notch = FALSE)

# Display the boxplot
plot_var

# Create a scatter plot of mean vs. variance of pycnidia
plot <- ggplot(sf.N.L_pair, aes(x=Mean, y=Var)) + 
  geom_point()+
  geom_text(x=30, y=2500, label="cor = 0.74 - pval < 2.2e-16 \n cov = 4804", col="black", size = 5) +
  geom_smooth(color="darkred", fill="red")+
  xlab("Average pycnidia surface of focal plants")+
  ylab("Pycnidia variance of focal plants")+
  theme(axis.text.x = element_text(color = "black", size = 20, face = "bold",angle = 0, hjust = ),
  axis.title.x = element_text(color = "black", size = 20, face = "bold"), 
  strip.text.x = element_text(color = "black",size=20, face = "bold"),
  axis.text.y = element_text(color = "black", size = 20, face = "bold",angle = 0, hjust = ),
  axis.title.y = element_text(color = "black", size = 20, face = "bold",angle = 90, hjust = ))

# Display the scatter plot
plot

# Save the scatter plot to a file
ggsave(file = "Corr_mean_var_DW_Pycnidia.svg", plot = plot, width = 7, height = 7)

```



# Load data for durum wheat - STB



```{r data - BW, results="hide"}

# Remove all objects from the current workspace
remove(list=ls())

# Set the working directory to the specified path
setwd("D:/Mes Donnees/LAURA/Redaction/1. Matrices/Matrices - v2/Data & script")

# Load the dataset from a CSV file
data_BW <- read.table("Bread_wheat_STB.csv", sep = ";", dec = ",", head = TRUE)

# Convert specific columns to appropriate data types
data_BW$Focal <- as.factor(data_BW$Focal)         # Convert 'Focal' to a factor
data_BW$Neighbor <- as.factor(data_BW$Neighbor)   # Convert 'Neighbor' to a factor
data_BW$Pair <- as.factor(data_BW$Pair)           # Convert 'Pair' to a factor
data_BW$Condition <- as.factor(data_BW$Condition) # Convert 'Condition' to a factor
data_BW$Tray <- as.factor(data_BW$Tray)           # Convert 'Tray' to a factor
data_BW$N <- as.numeric(data_BW$N)                # Convert 'N' to numeric
data_BW$PN <- as.numeric(data_BW$PN)              # Convert 'PN' to numeric

# Display the structure of the dataset to verify the changes
str(data_BW)

```


# Necrosis analysis


# Mean pairwise comparisons

```{r mean pairwise comparisons - BW - N}

# Fit an aligned rank transform (ART) model
model = art(N ~ (1|Tray) + Pair, data = data_BW)

# Perform ANOVA on the ART model
anova(model)

# Conduct pairwise comparisons using the ART model
sign <- art.con(model, "Pair", adjust = "none")

# Increase the maximum number of printed lines to display all results
options(max.print=5000)

# Convert the pairwise comparisons to a data frame
sign_table <- data.frame(sign)

# Apply Bonferroni correction: filter results with p-value < 0.00909 (0.1/11)
sign_table2<-sign_table[sign_table$p.value<0.009090909,] 

# Annotated pairwise comparisons with significant results after Bonferroni correction:
# Acc_Acc - Acc_Rub: 8.005278e-03
# Ber_Acc - Ber_Ber: 1.037232e-03
# Ber_Ber - Ber_Cen: 7.463585e-03
# Ber_Ber - Ber_SYM: 2.296271e-03
# Cen_Cen - Cen_SYM: 1.397024e-03
# Des_Des - Des_Gen: 1.848354e-03
# Fru_Cen - Fru_Fru: 7.830837e-04
# Gha_Gha - Gha_Rec: 1.300476e-03
# Rec_Cen - Rec_Rec: 3.347074e-04
# SYM_Fru - SYM_SYM: 1.392333e-03
# Tit_Ber - Tit_Tit: 3.136388e-03
# Tit_Cen - Tit_Tit: 1.398864e-03

# 12 mixtures = 9.09% mixtures showing NMS

```

## Create mean heatmap for pairwise comparisons

```{r mean pairwise heatmap - BW - N, include=FALSE}

# Generate estimated marginal means (least squares means)
emm <- emmeans(artlm(model, "Pair"), ~ Pair, pbktest.limit = 200000)
leastsquare <- as.data.frame(emm)

# Merge the estimated marginal means with the original dataset
data_BW_2 <- merge(leastsquare, data_BW, by = "Pair")

# Initialize a data frame to store the results
Pair=unique(data_BW_2$Pair)
Paires<-data.frame()
compt=1

# Calculate the mean 'emmean' for each pair
for(i in Pair){ 
  sub<-subset(data_BW_2,Pair==i)
  x<-mean(sub$emmean)
  print(x)
  Paires[compt,1]<-i
  Paires[compt,2]<-x
  f<-unique(sub$Focal)
  n<-unique(sub$Neighbor)
  Paires[compt,3]<-f
  Paires[compt,4]<-n
  compt<-compt+1
}

# Rename the columns of the data frame
colnames(Paires)<-c('Pair','emmean', 'Focal', 'Neighbor')

# Initialize data frames for storing normalized values and pairs
Focal=unique(Paires$Focal)
liste<-data.frame(nrow=1:12)
liste$nrow<-as.character(liste$nrow)
liste_pair<-data.frame(nrow=1:12)
liste_pair$nrow<-as.character(liste_pair$nrow)

# Normalize the emmeans by the mean of each focal pair
for(i in Focal){
  pair_sub<-subset(Paires,Focal==i)
  
  pair_sub_ii<-subset(pair_sub,Neighbor==i)
  x<-pair_sub_ii$emmean
  print(x)
  
  y<-pair_sub$emmean
  print(y)
  norm<-pair_sub$emmean/x
  norm<-as.data.frame(norm)
  norm[,2]<-row.names(norm)
  colnames(norm)<-c('norm', 'nrow')
  print(norm)
  liste <- liste %>% full_join(norm, by="nrow")
  
  pair<-pair_sub$Pair
  pair<-as.data.frame(pair)
  pair[,2]<-row.names(pair)
  colnames(pair)<-c('Pair', 'nrow')
  liste_pair <- liste_pair %>% full_join(pair, by="nrow")
  
}

# Process the normalized values and pairs
liste<-liste[,2:13]
colnames(liste)<-1:12
list<-as.list(liste)
list_num<-data.frame(unlist(list))
list_ssNA<-subset(list_num, unlist.list.!="NA")

liste_pair<-liste_pair[,2:13]
colnames(liste_pair)<-1:12
list_pair<-as.list(liste_pair)
list_pair<-data.frame(unlist(list_pair))
list_pair_ssNA<-subset(list_pair, unlist.list_pair.!="NA")

# Merge the processed values with the original pairs
y_pur<-merge(list_ssNA, list_pair_ssNA, by=0)
colnames(y_pur)<-c('nrow','y_pur','Pair')
Pair<-merge(y_pur, Paires, by = "Pair")


# Reorder the levels of the 'Focal' factor
Pair$Focal <- factor(Pair$Focal,levels=rev(unique(Paires$Focal)))

# Create a mean heatmap using ggplot2
heatmap_y_mix_pure_mean_pure<-ggplot(Pair,aes(x=Neighbor,y=Focal)) + 
  geom_tile(aes(fill = y_pur), colour = "black", size = 1)+
  scale_fill_gradient2(low = "darkblue", high = "red", mid = "white", midpoint = 1)+
  theme_minimal()+
  geom_text(x=10, y=12.3, label=".", col="black", size = 10)+
  geom_text(x=1, y=9.8, label="*", col="black", size = 10)+
  geom_text(x=4, y=10.3, label=".", col="black", size = 10)+
  geom_text(x=11, y=9.8, label="*", col="black", size = 10)+
  geom_text(x=11, y=8.8, label="*", col="black", size = 10)+
  geom_text(x=7, y=7.8, label="*", col="black", size = 10)+
  geom_text(x=4, y=6.8, label="**", col="black", size = 10)+
  geom_text(x=9, y=4.8, label="*", col="black", size = 10)+
  geom_text(x=4, y=3.8, label="**", col="black", size = 10)+
  geom_text(x=6, y=1.8, label="*", col="black", size = 10)+
  geom_text(x=3, y=0.8, label="*", col="black", size = 10)+
  geom_text(x=4, y=0.8, label="*", col="black", size = 10)+
  ylab("Focal") +
  xlab("Neighbor")+
  ggtitle("Necrosis surface / leaf surface of the focal plant (N/L)")+
  labs(fill="Mean(Mix) / Mean(Pure)")+
  theme(plot.title =element_text(color = "black", size = 15, face = "italic",hjust =0.5),#mise en forme
        legend.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 12, face = "bold",angle = 90, hjust = ),
        axis.title.x = element_text(color = "black", size = 12, face = "bold"),
        strip.text.x = element_text(color = "black",size=12, face = "bold"),
        axis.text.y = element_text(color = "black", size = 12, face = "bold",angle = 0, hjust = ),
        axis.title.y = element_text(color = "black", size = 12, face = "bold",angle = 90, hjust = )
  )

# Display the heatmap
heatmap_y_mix_pure_mean_pure

# Save the heatmap as an SVG file
ggsave(file="Matrice_NonParam_BW_Necrosis.svg", plot=heatmap_y_mix_pure_mean_pure, width=9, height=9)

```

## Variance pairwise comparisons

```{r variance pairwise comparisons - BW - N}

# Initialize variables
Pair = unique(data_BW$Pair)
Paires = data.frame()
compt = 1

# Calculate variances for each pair and store in Paires dataframe
for(i in Pair){ 
  sub<-subset(data_BW,Pair==i)
  x<-var(sub$N)
  print(x)
  Paires[compt,1]<-i
  Paires[compt,2]<-x
  f<-unique(sub$Focal)
  n<-unique(sub$Neighbor)
  Paires[compt,3]<-f
  Paires[compt,4]<-n
  compt<-compt+1
   }

# Rename columns for clarity
colnames(Paires)<-c('Pair','var', 'Focal', 'Neighbor')

# Initialize dataframes for normalized values and pairs
Focal=unique(Paires$Focal)
liste<-data.frame(nrow=1:12)
liste$nrow<-as.character(liste$nrow)
liste_pair<-data.frame(nrow=1:12)
liste_pair$nrow<-as.character(liste_pair$nrow)

# Normalize the variances by the mean of each focal pair
for(i in Focal){
  pair_sub<-subset(Paires, Focal==i)
  
  pair_sub_ii<-subset(pair_sub, Neighbor==i)
  x<-pair_sub_ii$var
  print(x)
  
  y<-pair_sub$var
  print(y)
  norm<-pair_sub$var/x
  norm<-as.data.frame(norm)
  norm[,2]<-row.names(norm)
  colnames(norm)<-c('norm', 'nrow')
  print(norm)
  liste <- liste %>% full_join(norm, by="nrow")
  
  pair<-pair_sub$Pair
  pair<-as.data.frame(pair)
  pair[,2]<-row.names(pair)
  colnames(pair)<-c('Pair', 'nrow')
  liste_pair <- liste_pair %>% full_join(pair, by="nrow")
  
}

# Process the normalized values and pairs
liste<-liste[,2:13]
colnames(liste)<-1:12
list<-as.list(liste)
list_num<-data.frame(unlist(list))
list_ssNA<-subset(list_num, unlist.list.!="NA")

liste_pair<-liste_pair[,2:13]
colnames(liste_pair)<-1:12
list_pair<-as.list(liste_pair)
list_pair<-data.frame(unlist(list_pair))
list_pair_ssNA<-subset(list_pair, unlist.list_pair.!="NA")

# Merge the processed values with the original pairs
y_pur<-merge(list_ssNA, list_pair_ssNA, by=0)
colnames(y_pur)<-c('nrow','y_pur','Pair')
Pair<-merge(y_pur, Paires, by = "Pair")

# Perform pairwise variance tests
pvals_var_all <- data.frame()

Focal=unique(Paires$Focal)
Neighbor=unique(Paires$Neighbor)

for(i in Focal){
  pair_sub<-subset(data_BW,Focal==i)
  
  for(j in Neighbor){
    # Ajouter la condition pour exécuter la boucle uniquement si j est différent de i
    if (j != i) {
      pair_sub_ii <- subset(pair_sub, Neighbor == i | Neighbor == j)
      
      levene <- leveneTest(N ~ Condition, data = pair_sub_ii)
      fligner <- fligner.test(N ~ Condition, data = pair_sub_ii)

      pvals_var <- data.frame()
      pvals_var[1, 1] <- i
      pvals_var[1, 2] <- j
      pvals_var[1, 3] <- levene$`Pr(>F)`[1]
      pvals_var[1, 4] <- fligner$p.value

      colnames(pvals_var) <- c('Focal', 'Neighbor', 'pval_Levene', 'pval_Fligner')
      pvals_var_all <- rbind(pvals_var_all, pvals_var)
    }
  }
}

# Adjust p-values for multiple comparisons (Bonferroni correction)
pvals_var_all$pval_Levene <- pvals_var_all$pval_Levene*11
pvals_var_all$pval_Fligner <- pvals_var_all$pval_Fligner*11
pvals_var_all <- pvals_var_all[order(pvals_var_all$pval_Fligner), ]

# Annotated pairwise comparisons with significant results after Bonferroni correction:
# Focal     Neighbor  pval_Levene   pval_Fligner  
# SYMattis  Bergamo   7.566479e-09  8.028049e-05  ***
# Bergamo   Accroc    1.183542e-01  1.299598e-04  ***
# Bergamo   SYMattis  1.402896e-01  2.002074e-04  ***
# Descartes Ghayta    1.233956e-04  5.404730e-04  ***
# Geny (AB) Accroc    2.846591e-02  5.495394e-04  ***
# Accroc    Bergamo   3.266480e-03  7.862811e-04  ***
# SYMattis  Ghayta    3.349031e-06  1.257707e-03  **
# Accroc    Geny (AB) 1.052518e-02  4.858227e-03  **
# SYMattis  Geny (AB) 2.217538e-04  9.040999e-03  **
# Bergamo   Geny (AB) 8.426341e-01  1.099151e-02  *
# Rubisko   Geny (AB) 2.276796e-01  1.134989e-02  *
# Centurion Titlis    1.502621e-01  1.233881e-02  *
# Titlis    Descartes 2.553012e-02  1.534845e-02  *
# Ghayta    Rubisko   8.252292e-01  1.585896e-02  *
# SYMattis  Descartes 1.028085e-03  1.613225e-02  *
# Geny (AB) Centurion 6.473045e-01  4.145409e-02  *
# Centurion SYMattis  1.118138e+00  4.190798e-02  *
# Fructidor Centurion 1.949513e-01  4.215513e-02  *
# Titlis    SYMattis  3.383048e-02  4.497647e-02  *
# Recital   Rubisko   1.875513e-01  4.650920e-02  *
# SYMattis  Centurion 8.198798e-04  5.627800e-02  .
# Centurion Geny (AB) 8.960032e-01  6.211341e-02  .
# Rubisko   Recital   2.906700e-01  8.642936e-02  .
# Fructidor Descartes 1.952299e-01  8.846788e-02  .
# SYMattis  Apache    1.640033e-02  9.077468e-02  .

# 25 mixtures (18.94% of mixtures) have a variance different from pure

```

## Variance heatmap of pairwise comparisons

```{r mean pairwise heatmap - BW - N , include=FALSE}

# Replace the values > 4 by NA
Pair$y_pur <- ifelse(Pair$y_pur > 4, NA, Pair$y_pur)

# Reorder the levels of the 'Focal' factor
Pair$Focal <- factor(Pair$Focal, levels = rev(sort(unique(Pair$Focal))))

# Create a heatmap using ggplot2
heatmap_y_mix_pure_var <- ggplot(Pair,aes(x=Neighbor,y= Focal)) + 
  geom_tile(aes(fill = y_pur), colour = "black", size = 1)+
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "red", midpoint = 1, na.value = "#A52A2A") +
  theme_minimal()+
  geom_text(x=2, y=2.3, label=".", col="black", size = 10)+
  geom_text(x=5, y=7.3, label=".", col="black", size = 10)+
  geom_text(x=9, y=3.3, label=".", col="black", size = 10)+
  geom_text(x=7, y=9.3, label=".", col="black", size = 10)+
  geom_text(x=4, y=2.3, label=".", col="black", size = 10)+
  geom_text(x=7, y=9.8, label="*", col="black", size = 10)+
  geom_text(x=7, y=2.8, label="*", col="black", size = 10)+
  geom_text(x=12, y=8.8, label="*", col="black", size = 10)+
  geom_text(x=5, y=0.8, label="*", col="black", size = 10)+
  geom_text(x=10, y=4.8, label="*", col="black", size = 10)+
  geom_text(x=5, y=1.8, label="*", col="black", size = 10)+
  geom_text(x=4, y=5.8, label="*", col="black", size = 10)+
  geom_text(x=11, y=8.8, label="*", col="black", size = 10)+
  geom_text(x=4, y=6.8, label="*", col="black", size = 10)+
  geom_text(x=11, y=0.8, label="*", col="black", size = 10)+
  geom_text(x=10, y=3.8, label="*", col="black", size = 10)+
  geom_text(x=8, y=1.8, label="**", col="black", size = 10)+
  geom_text(x=7, y=11.8, label="**", col="black", size = 10)+
  geom_text(x=7, y=1.8, label="**", col="black", size = 10)+
  geom_text(x=3, y=1.8, label="***", col="black", size = 10)+
  geom_text(x=1, y=9.8, label="***", col="black", size = 10)+
  geom_text(x=11, y=9.8, label="***", col="black", size = 10)+
  geom_text(x=8, y=7.8, label="***", col="black", size = 10)+
  geom_text(x=1, y=5.8, label="***", col="black", size = 10)+
  geom_text(x=3, y=11.8, label="***", col="black", size = 10)+
  
  ylab("Focal") +
  xlab("Neighbor")+
  ggtitle("Necrosis surface / leaf surface of the focal plant (N/L)")+
  labs(fill="Var(Mix) / Var(Pure)")+
  theme(plot.title =element_text(color = "black", size = 15, face = "italic",hjust =0.5),#mise en forme
        legend.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 12, face = "bold",angle = 90, hjust = ),
        axis.title.x = element_text(color = "black", size = 12, face = "bold"),
        strip.text.x = element_text(color = "black",size=12, face = "bold"),
        axis.text.y = element_text(color = "black", size = 12, face = "bold",angle = 0, hjust = ),
        axis.title.y = element_text(color = "black", size = 12, face = "bold",angle = 90, hjust = )
  )

# Display the heatmap
heatmap_y_mix_pure_var

# Save the heatmap as an SVG file
ggsave(file = "Matrice_NonParam_var_BW_Necrosis.svg", plot = heatmap_y_mix_pure_var, width = 9, height = 9)


```

## Variance explained by genetic effects

```{r DGE-IGE - BW - N, include=FALSE}

# Fit the aligned rank transform model to the data
model = art(N ~ Focal * Neighbor, data = data_BW_2)

# Perform ANOVA on the fitted model
Result = anova(model)

# Calculate partial eta squared for each term in the ANOVA table
Result$part.eta.sq = with(Result, `Sum Sq` / (`Sum Sq` + `Sum Sq.res`))

# Display the ANOVA result with partial eta squared
Result

# Perform ANOVA on the fitted model again for redundancy (to show another way to calculate partial eta squared)
Result2 = anova(model)

# Calculate partial eta squared using an alternative method
Result2$part.eta.sq = with(Result2, `F value` * `Df` / (`F value` * `Df` + `Df.res`))

# Display the second ANOVA result with partial eta squared
Result2

```

## Boxplots mix-pur with variances & means

```{r mean - variances - BW - N}

# Conduct Wilcoxon test & fligner test for N by condition
wilcox.test(N~Condition, data=data_BW)
fligner.test(N~Condition, data=data_BW)

# Calculate variances for mix and pure conditions
Mix <- subset(data_BW, data_BW$Condition == "Mix")
Mix <- Mix$N
var(Mix)
Pure <- subset(data_BW, data_BW$Condition == "Pure")
Pure <- Pure$N
var(Pure)

# Plotting
violin <- ggplot(data_BW_2, aes(x=factor(Condition, level=c("Pure","Mix")), y=N)) + geom_violin(trim=FALSE)
plot <- violin + geom_boxplot(width=0.3, col="#9cb264") + stat_summary(fun=mean, geom="point", shape=23, size=5, color="#7a3776")+
  geom_text(x=1.5, y=1.45, label="*", col="#7a3776", size = 20)+
  geom_text(x=1, y=1.45, label="μ(pure) = 0.34", col="#7a3776", size = 10)+
  geom_text(x=2, y=1.45, label="μ(mix) = 0.30", col="#7a3776", size = 10)+
  geom_text(x=1.5, y=1.35, label="**", col="#9cb264", size = 20)+
  geom_text(x=1, y=1.35, label="var(pure) = 0.15", col="#9cb264", size = 10)+
  geom_text(x=2, y=1.35, label="var(mix) = 0.14", col="#9cb264", size = 10)+
  ylab("Necrosis surface / leaf surface (N/L)") +
  ggtitle("Bread wheat")+
  theme(plot.title =element_text(color = "black", size = 40, face = "italic",hjust =0.5),
        axis.text.x = element_text(color = "black", size = 40, face = "bold",angle = 0, hjust = ),
  axis.title.x = element_text(color = "black", size = 40, face = "bold"), 
  strip.text.x = element_text(color = "black",size=40, face = "bold"),
  axis.text.y = element_text(color = "black", size = 40, face = "bold",angle = 0, hjust = ),
  axis.title.y = element_text(color = "black", size = 40, face = "bold",angle = 90, hjust = ))
plot <- plot + ylim(-0.25, 1.5)
print(plot)

# Save the plot as an SVG file
ggsave(file="MixPur_brut_mean_var_BW_Necrosis.svg", plot=plot, width=15, height=15)

```

## Variances vs. mean analysis 

```{r var ~ mean - BW - N, include=FALSE}

# Initialize variables and data frame
pair = unique(data_BW$Pair)
sf.N.L_pair <- data.frame()
n = 1

# Loop through each modality to calculate variance and mean of necrosis surface
for(i in pair){
  symptom=subset(data_BW,Pair==i)
  v<-var(symptom$N)
  m<-mean(symptom$N)
  print(v)
  print(m)
  new<-c(i, v, m)
  print(new)
  sf.N.L_pair[nrow(sf.N.L_pair)+1,1:3]<-c(i, v, m)
  n=n+1
}

# Rename columns
colnames(sf.N.L_pair)<-c('Pair', 'Var', 'Mean')
sf.N.L_pair[,2]<-as.numeric(sf.N.L_pair[,2])
sf.N.L_pair[,3]<-as.numeric(sf.N.L_pair[,3])
sf.N.L_pair<-sf.N.L_pair[order(sf.N.L_pair[,3]),]

# Merge the calculated variance and mean back to the original data
data_BW_3<-merge(data_BW,sf.N.L_pair)
data_BW_3<-data_BW_3[order(data_BW_3[,9]),]

# Create a boxplot of necrosis surface by pair, ordered by mean
plot_var <- ggplot(data_BW_3, aes(x=reorder(Pair,Mean), y=N)) + 
  geom_boxplot(fill="red",outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE)

# Display the boxplot
plot_var

# Create a scatter plot of mean vs. variance of necrosis surface
plot <- ggplot(sf.N.L_pair, aes(x=Mean, y=Var)) + 
  geom_point()+
  geom_smooth(color="darkred", fill="red")+
  theme(axis.text.x = element_text(color = "black", size = 20, face = "bold",angle = 0, hjust = ),
  axis.title.x = element_text(color = "black", size = 20, face = "bold"), 
  strip.text.x = element_text(color = "black",size=20, face = "bold"),
  axis.text.y = element_text(color = "black", size = 20, face = "bold",angle = 0, hjust = ),
  axis.title.y = element_text(color = "black", size = 20, face = "bold",angle = 90, hjust = ))

# Display the scatter plot
plot

# Save the scatter plot to a file
ggsave(file="Corr_mean_var_BW_Necrosis.svg", plot=plot, width=7, height=7)

```


# Pycnidia analysis


## Mean pairwise comparisons

```{r mean pairwise - BW - P}

# Fit an aligned rank transform (ART) model
model = art(PN ~ (1|Tray) + Pair, data = data_BW)

# Perform ANOVA on the ART model
anova(model)

# Conduct pairwise comparisons using the ART model
sign <- art.con(model, "Pair", adjust = "none")

# Increase the maximum number of printed lines to display all results
options(max.print=5000)

# Convert the pairwise comparisons to a data frame
sign_table <- data.frame(sign)

# Apply Bonferroni correction: filter results with p-value < 0.00909 (0.1/11)
sign_table2<-sign_table[sign_table$p.value<0.009090909,] 

# Annotated pairwise comparisons with significant results after Bonferroni correction:
# Cen_Cen - Cen_Tit: 4.654860e-03
# Des_Des - Des_Tit: 6.507255e-03

# 3 mixtures = 1.6% of mixtures showing significant differences

```

## Create mean heatmap for pairwise comparisons

```{r mean pairwise heatmap - BW - P, include=FALSE}

# Generate estimated marginal means (least squares means)
emm <- emmeans(artlm(model, "Pair"), ~ Pair, pbktest.limit = 200000)

# Convert the estimated marginal means to a data frame
leastsquare <- as.data.frame(emm)

# Merge the estimated marginal means with the original dataset
data_BW_2 <- merge(leastsquare, data_BW, by = "Pair")

# Initialize a data frame to store the results
Pair <- unique(data_BW_2$Pair)
Paires <- data.frame()
compt <- 1

# Calculate the mean 'emmean' for each pair
for(i in Pair) {
  sub <- subset(data_BW_2, Pair == i)
  x <- mean(sub$emmean)
  Paires[compt, 1] <- i
  Paires[compt, 2] <- x
  f <- unique(sub$Focal)
  n <- unique(sub$Neighbor)
  Paires[compt, 3] <- f
  Paires[compt, 4] <- n
  compt <- compt + 1
}

# Rename the columns of the data frame
colnames(Paires) <- c('Pair', 'emmean', 'Focal', 'Neighbor')

# Initialize data frames for storing normalized values and pairs
Focal=unique(Paires$Focal)
liste<-data.frame(nrow=1:12)
liste$nrow<-as.character(liste$nrow)
liste_pair<-data.frame(nrow=1:12)
liste_pair$nrow<-as.character(liste_pair$nrow)

# Normalize the emmeans by the mean of each focal pair
for(i in Focal){
  pair_sub<-subset(Paires,Focal==i)
  
  pair_sub_ii<-subset(pair_sub,Neighbor==i)
  x<-pair_sub_ii$emmean
  print(x)
  
  y<-pair_sub$emmean
  print(y)
  norm<-pair_sub$emmean/x
  norm<-as.data.frame(norm)
  norm[,2]<-row.names(norm)
  colnames(norm)<-c('norm', 'nrow')
  print(norm)
  liste <- liste %>% full_join(norm, by="nrow")
  
  pair<-pair_sub$Pair
  pair<-as.data.frame(pair)
  pair[,2]<-row.names(pair)
  colnames(pair)<-c('Pair', 'nrow')
  liste_pair <- liste_pair %>% full_join(pair, by="nrow")
  
}

# Process the normalized values and pairs
liste<-liste[,2:13]
colnames(liste)<-1:12
list<-as.list(liste)
list_num<-data.frame(unlist(list))
list_ssNA<-subset(list_num, unlist.list.!="NA")

liste_pair<-liste_pair[,2:13]
colnames(liste_pair)<-1:12
list_pair<-as.list(liste_pair)
list_pair<-data.frame(unlist(list_pair))
list_pair_ssNA<-subset(list_pair, unlist.list_pair.!="NA")

# Merge the processed values with the original pairs
y_pur<-merge(list_ssNA, list_pair_ssNA, by=0)
colnames(y_pur)<-c('nrow','y_pur','Pair')
Pair<-merge(y_pur, Paires, by = "Pair")

# Reorder the levels of the 'Focal' factor
Pair$Focal <- factor(Pair$Focal,levels=rev(unique(Paires$Focal)))

# Create a mean heatmap using ggplot2
heatmap_y_mix_pure_mean_pure<-ggplot(Pair,aes(x=Neighbor,y=Focal)) + 
  geom_tile(aes(fill = y_pur), colour = "black", size = 1)+
  scale_fill_gradient2(low = "darkblue", high = "red", mid = "white", midpoint = 1)+
  theme_minimal()+
  geom_text(x=12, y=8.3, label=".", col="black", size = 10) +
  geom_text(x=12, y=9.3, label=".", col="black", size = 10) +
  # ylab("Focal") +
  xlab("Neighbor")+
  ggtitle("Pycnidia surface / necrosis surface of the focal plant (P/N)")+
  labs(fill="Mean(Mix) / Mean(Pure)")+
  theme(plot.title =element_text(color = "black", size = 15, face = "italic",hjust =0.5),#mise en forme
        legend.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 12, face = "bold",angle = 90, hjust = ),
        axis.title.x = element_text(color = "black", size = 12, face = "bold"),
        strip.text.x = element_text(color = "black",size=12, face = "bold"),
        axis.text.y = element_text(color = "black", size = 12, face = "bold",angle = 0, hjust = ),
        axis.title.y = element_text(color = "black", size = 12, face = "bold",angle = 90, hjust = )
  )

# Display the heatmap
heatmap_y_mix_pure_mean_pure

# Save the heatmap as SVG file
ggsave(file="Matrice_NonParam_BW_Pycnidia.svg", plot=heatmap_y_mix_pure_mean_pure, width=9, height=9)

```

## Variance paiwise comparisons 

```{r variance paiwise comparisons - BW - P}

# Initialize variables
Pair <- unique(data_BW$Pair)
Paires <- data.frame()
compt <- 1

# Calculate variances for each pair and store in Paires dataframe
for (i in Pair) { 
  sub <- subset(data_BW, Pair == i)
  x <- var(sub$PN)
  print(x)
  Paires[compt, 1] <- i
  Paires[compt, 2] <- x
  f <- unique(sub$Focal)
  n <- unique(sub$Neighbor)
  Paires[compt, 3] <- f
  Paires[compt, 4] <- n
  compt <- compt + 1
}

# Rename columns for clarity
colnames(Paires) <- c('Pair', 'var', 'Focal', 'Neighbor')

# Initialize dataframes for normalized values and pairs
Focal=unique(Paires$Focal)
liste<-data.frame(nrow=1:12)
liste$nrow<-as.character(liste$nrow)
liste_pair<-data.frame(nrow=1:12)
liste_pair$nrow<-as.character(liste_pair$nrow)

# Normalize the variances by the mean of each focal pair
for(i in Focal){
  pair_sub<-subset(Paires, Focal==i)
  
  pair_sub_ii<-subset(pair_sub, Neighbor==i)
  x<-pair_sub_ii$var
  print(x)
  
  y<-pair_sub$var
  print(y)
  norm<-pair_sub$var/x
  norm<-as.data.frame(norm)
  norm[,2]<-row.names(norm)
  colnames(norm)<-c('norm', 'nrow')
  print(norm)
  liste <- liste %>% full_join(norm, by="nrow")
  
  pair<-pair_sub$Pair
  pair<-as.data.frame(pair)
  pair[,2]<-row.names(pair)
  colnames(pair)<-c('Pair', 'nrow')
  liste_pair <- liste_pair %>% full_join(pair, by="nrow")
  
}

# Process the normalized values and pairs
liste<-liste[,2:13]
colnames(liste)<-1:12
list<-as.list(liste)
list_num<-data.frame(unlist(list))
list_ssNA<-subset(list_num, unlist.list.!="NA")

liste_pair<-liste_pair[,2:13]
colnames(liste_pair)<-1:12
list_pair<-as.list(liste_pair)
list_pair<-data.frame(unlist(list_pair))
list_pair_ssNA<-subset(list_pair, unlist.list_pair.!="NA")

# Merge the processed values with the original pairs
y_pur<-merge(list_ssNA, list_pair_ssNA, by=0)
colnames(y_pur)<-c('nrow','y_pur','Pair')
Pair<-merge(y_pur, Paires, by = "Pair")

# Perform pairwise variance tests
pvals_var_all <- data.frame()

Focal=unique(Paires$Focal)
Neighbor=unique(Paires$Neighbor)

for(i in Focal){
  pair_sub<-subset(data_BW,Focal==i)
  
  for(j in Neighbor){
    # Ajouter la condition pour exécuter la boucle uniquement si j est différent de i
    if (j != i) {
      pair_sub_ii <- subset(pair_sub, Neighbor == i | Neighbor == j)
      
      levene <- leveneTest(PN ~ Condition, data = pair_sub_ii)
      fligner <- fligner.test(PN ~ Condition, data = pair_sub_ii)

      pvals_var <- data.frame()
      pvals_var[1, 1] <- i
      pvals_var[1, 2] <- j
      pvals_var[1, 3] <- levene$`Pr(>F)`[1]
      pvals_var[1, 4] <- fligner$p.value

      colnames(pvals_var) <- c('Focal', 'Neighbor', 'pval_Levene', 'pval_Fligner')
      pvals_var_all <- rbind(pvals_var_all, pvals_var)
    }
  }
}

# Adjust p-values for multiple comparisons (Bonferroni correction)
pvals_var_all$pval_Levene <- pvals_var_all$pval_Levene*13
pvals_var_all$pval_Fligner <- pvals_var_all$pval_Fligner*13
pvals_var_all <- pvals_var_all[order(pvals_var_all$pval_Fligner), ]
pvals_var_all

# Annotated pairwise comparisons with significant results after Bonferroni correction:
#   Focal     Neighbour   pval_Levene   pval_Fligner  
#   Centurion Titlis      0.002997090   9.310193e-05  ***
#   Titlis    Descartes   0.001357412   2.335945e-03  **
#   Geny (AB) Centurion   1.354170718   3.928537e-03  **
#   Titlis    Centurion   0.004448451   4.371542e-03  **
#   Titlis    Accroc      0.005152907   4.715933e-03  **
#   Titlis    Bergamo     0.008069348   9.224496e-03  **
#   Descartes Bergamo     2.160216433   9.908181e-03  **
#   Geny (AB) Accroc      1.420734685   1.164669e-02  *
#   Centurion Fructidor   0.075305654   1.379039e-02  *
#   Titlis    Apache      0.015179382   1.515646e-02  *
#   Centurion Bergamo     0.064121709   1.654414e-02  *
#   Centurion Geny (AB)   0.115770996   1.735435e-02  *
#   Fructidor Titlis      0.308019944   1.869588e-02  *
#   Titlis    Geny (AB)   0.032527446   2.004132e-02  *
#   Bergamo   Descartes   2.276633520   2.143306e-02  *
#   Titlis    SYMattis    0.015504801   2.639702e-02  *
#   Bergamo   Rubisko     1.192856426   3.366814e-02  *
#   Geny (AB) Fructidor   11.217225384  3.899231e-02  *
#   Titlis    Fructidor   0.033336411   4.484693e-02  *
#   Descartes Recital     2.693449233   4.645531e-02  *
#   Accroc    Recital     0.012424249   5.403685e-02  .
#   Centurion Apache      1.347921235   7.056202e-02  .
#   Fructidor Centurion   2.652741732   7.554338e-02  .
#   Descartes Rubisko     2.716428216   8.728595e-02  .

# 24 mixtures (17.42% of mixtures) have a variance different from pure

```

## Variance heatmap of pairwise comparisons

```{r variance pairwise heatmap - BW - P, include=FALSE}

# Replace values > 4 with NA
Pair$y_pur <- ifelse(Pair$y_pur > 4, NA, Pair$y_pur)

# Reorder the levels of the 'Focal' factor
Pair$Focal <- factor(Pair$Focal, levels = rev(sort(unique(Pair$Focal))))

# Create a heatmap using ggplot2
heatmap_y_mix_pure_var <- ggplot(Pair,aes(x=Neighbor,y= Focal)) + 
  geom_tile(aes(fill = y_pur), colour = "black", size = 1)+
  scale_fill_gradient2(low = "darkblue", mid = "white", high = "red", midpoint = 1, na.value = "#A52A2A") +
  theme_minimal()+
  
  geom_text(x=9, y=12.3, label=".", col="black", size = 10)+
  geom_text(x=2, y=9.3, label=".", col="black", size = 10)+
  geom_text(x=4, y=7.3, label=".", col="black", size = 10)+
  geom_text(x=10, y=8.3, label=".", col="black", size = 10)+
  
  geom_text(x=2, y=0.8, label="*", col="black", size = 10)+
  geom_text(x=7, y=0.8, label="*", col="black", size = 10)+
  geom_text(x=6, y=0.8, label="*", col="black", size = 10)+
  geom_text(x=11, y=0.8, label="*", col="black", size = 10)+
  geom_text(x=1, y=5.8, label="*", col="black", size = 10)+
  geom_text(x=6, y=5.8, label="*", col="black", size = 10)+
  geom_text(x=12, y=6.8, label="*", col="black", size = 10)+
  geom_text(x=9, y=7.8, label="*", col="black", size = 10)+
  geom_text(x=3, y=8.8, label="*", col="black", size = 10)+
  geom_text(x=7, y=8.8, label="*", col="black", size = 10)+
  geom_text(x=6, y=8.8, label="*", col="black", size = 10)+
  geom_text(x=5, y=9.8, label="*", col="black", size = 10)+
  geom_text(x=10, y=9.8, label="*", col="black", size = 10)+
  
  geom_text(x=3, y=7.8, label="**", col="black", size = 10)+
  geom_text(x=3, y=0.8, label="**", col="black", size = 10)+
  geom_text(x=1, y=0.8, label="**", col="black", size = 10)+
  geom_text(x=4, y=0.8, label="**", col="black", size = 10)+
  geom_text(x=5, y=0.8, label="**", col="black", size = 10)+
  geom_text(x=4, y=5.8, label="**", col="black", size = 10)+
  
  geom_text(x=12, y=8.8, label="***", col="black", size = 10)+
  
  ylab("Focal") +
  xlab("Neighbor")+
  ggtitle("Pycnidia surface / necrosis surface of the focal plant (P/N)")+
  labs(fill="Var(Mix) / Var(Pure)")+
  theme(plot.title =element_text(color = "black", size = 15, face = "italic",hjust =0.5),#mise en forme
        legend.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(color = "black", size = 12, face = "bold",angle = 90, hjust = ),
        axis.title.x = element_text(color = "black", size = 12, face = "bold"),
        strip.text.x = element_text(color = "black",size=12, face = "bold"),
        axis.text.y = element_text(color = "black", size = 12, face = "bold",angle = 0, hjust = ),
        axis.title.y = element_text(color = "black", size = 12, face = "bold",angle = 90, hjust = )
  )

# Display the heatmap
heatmap_y_mix_pure_var

# Save the heatmap as an SVG file
ggsave(file="Matrice_NonParam_var_BW_Pycnides.svg", plot=heatmap_y_mix_pure_var, width=9, height=9)

```

## Variance explained by genetic effects

```{r DGE-IGE - BW - P, include=FALSE}

# Fit the aligned rank transform model to the data
model = art(PN ~ Focal * Neighbor, data = data_BW_2)

# Perform ANOVA on the fitted model
Result = anova(model)

# Calculate partial eta squared for each term in the ANOVA table
Result$part.eta.sq = with(Result, `Sum Sq` / (`Sum Sq` + `Sum Sq.res`))

# Display the ANOVA result with partial eta squared
Result

# Perform ANOVA on the fitted model again for redundancy (to show another way to calculate partial eta squared)
Result2 = anova(model)

# Calculate partial eta squared using an alternative method
Result2$part.eta.sq = with(Result2, `F value` * `Df` / (`F value` * `Df` + `Df.res`))

# Display the second ANOVA result with partial eta squared
Result2

```

## Boxplots mix-pur with variances & means

```{r mean - variances - BW - P}

# Conduct Wilcoxon test & fligner test for N by condition
wilcox.test(PN~Condition, data=data_BW)
fligner.test(PN~Condition, data=data_BW)

# Calculate variances for mix and pure conditions
Mix <- subset(data_BW, data_BW$Condition == "Mix")
Mix <- Mix$PN
var(Mix)
Pure <- subset(data_BW, data_BW$Condition == "Pure")
Pure <- Pure$PN
var(Pure)

# Plotting
violin <- ggplot(data_BW_2, aes(x=factor(Condition, level=c("Pure","Mix")), y=PN)) + geom_violin(trim=FALSE)
plot <- violin + geom_boxplot(width=0.3, col="#9cb264") + stat_summary(fun=mean, geom="point", shape=23, size=5, color="#7a3776")+
  geom_text(x=1.5, y=0.058, label="ns", col="#7a3776", size = 10)+
  geom_text(x=1, y=0.058, label="μ(pure) = 0.027", col="#7a3776", size = 10)+
  geom_text(x=2, y=0.058, label="μ(mix) = 0.030", col="#7a3776", size = 10)+
  geom_text(x=1.5, y=0.053, label="*", col="#9cb264", size = 20)+
  geom_text(x=1, y=0.053, label="var(pure) = 0.002", col="#9cb264", size = 10)+
  geom_text(x=2, y=0.053, label="var(mix) = 0.005", col="#9cb264", size = 10)+
  ylab("Pycnidia surface / necrosis surface (P/N)")+
  ggtitle("Bread wheat")+
  theme(plot.title =element_text(color = "black", size = 40, face = "italic",hjust =0.5),
        axis.text.x = element_text(color = "black", size = 40, face = "bold",angle = 0, hjust = ),
  axis.title.x = element_text(color = "black", size = 40, face = "bold"), 
  strip.text.x = element_text(color = "black",size=40, face = "bold"),
  axis.text.y = element_text(color = "black", size = 40, face = "bold",angle = 0, hjust = ),
  axis.title.y = element_text(color = "black", size = 40, face = "bold",angle = 90, hjust = ))
plot <- plot + ylim(-0.02, 0.06)
print(plot)

# Save the plot as an SVG file
ggsave(file="MixPur_brut_mean_var_BW_Pycnidia.svg", plot=plot, width=15, height=15)

```

## Variances vs. mean analysis 

```{r var ~ mean - BW - P, include=FALSE}

# Initialize variables and data frame
pair = unique(data_BW$Pair)
sf.N.L_pair <- data.frame()
n = 1

# Loop through each modality to calculate variance and mean of necrosis surface
for (i in pair) {
  symptom = subset(data_BW, Pair == i)  # Subset data for the current pair
  v = var(symptom$PN)  # Calculate variance of necrosis surface
  m = mean(symptom$PN)  # Calculate mean of necrosis surface
  print(v)
  print(m)
  new = c(i, v, m)
  print(new)
  # Add the results to the data frame
  sf.N.L_pair[nrow(sf.N.L_pair) + 1, 1:3] <- c(i, v, m)
  n = n + 1
}

# Rename columns
colnames(sf.N.L_pair) <- c('Pair', 'Var', 'Mean')
sf.N.L_pair[,2] <- as.numeric(sf.N.L_pair[,2])
sf.N.L_pair[,3] <- as.numeric(sf.N.L_pair[,3])
# Order the data frame by mean
sf.N.L_pair <- sf.N.L_pair[order(sf.N.L_pair[,3]),]

# Merge the calculated variance and mean back to the original data
data_BW_3 <- merge(data_BW, sf.N.L_pair)
# Order the merged data by the calculated mean
data_BW_3 <- data_BW_3[order(data_BW_3[,9]),]

# Create a boxplot of pycnidia by pair, ordered by mean
plot_var <- ggplot(data_BW_3, aes(x=reorder(Pair,Mean), y=PN)) + 
  geom_boxplot(fill="red",outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE)

# Display the boxplot
plot_var

# Create a scatter plot of mean vs. variance of necrosis surface
plot <- ggplot(sf.N.L_pair, aes(x=Mean, y=Var)) + 
  geom_point()+
  geom_smooth(color="darkred", fill="red")+
  theme(axis.text.x = element_text(color = "black", size = 20, face = "bold",angle = 0, hjust = ),
  axis.title.x = element_text(color = "black", size = 20, face = "bold"), 
  strip.text.x = element_text(color = "black",size=20, face = "bold"),
  axis.text.y = element_text(color = "black", size = 20, face = "bold",angle = 0, hjust = ),
  axis.title.y = element_text(color = "black", size = 20, face = "bold",angle = 90, hjust = ))

# Display the scatter plot
plot

# Save the scatter plot to a file
ggsave(file="Corr_mean_var_BW_Pycnidia.svg", plot=plot, width=7, height=7)

```



